{RILS}-{ROLS} : Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-12-03 paper_resources/doc/bmc_jbd/bmc_jbd.tex
Abstract revision.

\begin{abstract} % abstract
In this paper, we solve the well-known symbolic regression problem. This problem has been intensively studied in the literature and has a wide range of applications. The symbolic regression models could provide insights towards establishing physical theory behind experimentally observed data. Practical applications include revealing complex ecological dynamics in ecology, discovering mutations effects on protein stability in biology, modeling analytic representations of the exciton binding energy in energy science, finding models for band gaps of NaCl-type compounds
in material science, just to name a few.

We propose an efficient meta-heuristic-based approach, called \textsc{RILS}-\textsc{ROLS} to solve this problem. \textsc{RILS}-\textsc{ROLS} is based on the following elements: ($i$) iterated local search is the method backbone, mainly solving combinatorial and some continuous aspects of the problem; ($ii$) ordinary least square method is focused on the continuous aspect of the search space -- it efficiently determines the best--fitting coefficients of linear combinations within solution equations;  ($iii$) a novel fitness function combines three important model quality measures: $R^2$ score, RMSE score, and the size of the model (or model complexity).

The experiments are conducted on the two well-known ground-truth benchmark sets called \textsc{Feynman} and \textsc{Strogatz}. \textsc{RILS}-\textsc{ROLS} was compared to 14 other competitors from the literature. Our method outperformed all 14 competitors with respect to the true symbolic model accuracy percentage under varying levels of noise. It also showed to be the most efficient considering the average running time of reaching the exact model.  
In addition to evaluation on known ground-truth datasets, a new randomly generated set of problem instances is introduced. The goal of \textsc{Random} dataset was to test the scalability of our method with respect to incremental equation sizes and different levels of noise. \textsc{RILS}-\textsc{ROLS} can be considered as a new state-of-the-art method for solving the problem of symbolic regression when ground-truth equations are known.
\end{abstract}

@@

\begin{keyword}
\kwd{Symbolic regression}
\kwd{Iterated local search}
\kwd{Ordinary least square}
\end{keyword}

>>>>

\begin{abstract} % abstract
In this paper, we solve the well-known symbolic regression problem that has been intensively studied in the literature and has a wide range of applications. 
% ACA: ove recenice vise prilice uvodu, tamo svakako i postoji, da ne bude preglomazan abstrakt
%The symbolic regression models could provide insights towards establishing physical theory behind experimentally observed data. 
%Practical applications include revealing complex ecological dynamics in ecology, discovering mutations effects on protein stability in biology, modeling analytic representations of the exciton binding energy in energy science, finding models for band gaps of NaCl-type compounds in material science, just to name a few.

We propose an efficient meta-heuristic-based approach, called \textsc{RILS}-\textsc{ROLS} to solve it. \textsc{RILS}-\textsc{ROLS} is based on the following two parts: ($i$) iterated local search serving as the method backbone, mainly solving combinatorial and some continuous aspects of the problem; ($ii$) ordinary least square method is focused on the continuous aspect of the search space -- it efficiently determines the best--fitting coefficients of linear combinations within solution equations. Beside that, we introduce a novel fitness function that combines three important model quality measures: $R^2$ score, RMSE score, and the size of the model (or model complexity), and carefully designed local search, which allows systematic search in proximity to candidate solution.

The experiments are conducted on the two well-known ground-truth benchmark sets called \textsc{Feynman} and \textsc{Strogatz}. \textsc{RILS}-\textsc{ROLS} was compared to 14 other competitors from the literature. Our method outperformed all 14 competitors with respect to the true symbolic model accuracy under varying levels of noise. It also showed to be the most efficient considering the average running time of reaching the exact model.  
In addition to evaluation on known ground-truth datasets, a new randomly generated set of problem instances is introduced. The goal of \textsc{Random} dataset was to test the scalability of our method with respect to incremental equation sizes and number of variables, under different levels of noise.

Statistical analysis of obtained experimental results confirmed that \textsc{RILS}-\textsc{ROLS} can be considered as a new state-of-the-art method for solving the problem of symbolic regression when ground-truth equations are known.
\end{abstract}

@@

\begin{keyword}
\kwd{symbolic regression}
\kwd{iterated local search}
\kwd{ordinary least square}
\end{keyword}
